---
title: "Untitled"
author: "Tejasav Ralli"
date: "2023-10-20"
output: pdf_document
---

## Question 1
```{r}
# Load the traffic data from the CSV file
traffic <- read.csv("traffic.csv", header = TRUE)

# Display the first few rows of the 'traffic' dataset
head(traffic)
```

### a)
```{r}
pairs(traffic, panel = panel.smooth)
cor(traffic)

#There are no clear associations between the response variable (spi) and the predictors 
#weather and fuel, however there is a weak linear negative link with the predictor road, 
#a weak negative linear relationship with the predictor transportation, 
#and a moderate positive linear relationship with the predictor weather.

#The predictors themselves don't appear to be related to one another. 

```

### b) Fit the full model 
```{r}
Model1 <- lm(spi ~ transport + road + weather + fuel + wind, data = traffic)
summary(Model1)
#Producing 95% conf. interval 
summary.Model1 <- summary(Model1)
se <- sqrt(diag(summary.Model1$cov.unscaled * summary.Model1$sigma^2)) [3]
qt(0.975, 56)
```


$$\hat{\beta}_{\text{weather}} +   t_{(n-p, 1-\alpha/2)} \times \text{se}\times {\hat{\beta}_{\text{weather}}}$$


$$\hat{\beta}_{\text{weather}} + t _{(56,0.975)} \times \text{s.e.}\times {\beta_{\text{weather}}}$$
$= 4.2456 + 2.003241 (0.4473)$
$= 5.141649699$

$$\hat{\beta}_{\text{weather}} -   t_{(n-p, 1-\alpha/2)} \times \text{se}\times {\hat{\beta}_{\text{weather}}}$$

$$\hat{\beta}_{\text{weather}} - t _{(56,0.975)} \times \text{s.e.}\times {\beta_{\text{weather}}}$$

$= 4.2456 - 2.003241 (0.4473)$
$= 3.349550301$

Conf. interval -> 5.141649699, 3.349550301


Comment - That is, we are 95% confident that for every percentage increase in relative weather, the spi will increase between 
$3.349550301$ and $5.141649699$ weather severity index on average.

### c) Conducting f-test

Theoretical Model

$$ \gamma_i = B_0 + B_1X_{i1} + B_2X_{i2} + B_3X_{i3} + B_4X_{i4} + \epsilon _i, \quad i = 1,2,\ldots,5 $$

Y is the response variable spi 
$X_{ij}$ are the predictor variables for the $i$-th observation.

\begin{enumerate}
\item $X_{i1}$ = annual mean transport of test locations.
\item $X_{i2}$ = annual mean road of test locations.
\item $X_{i3}$ = annual mean weather of test locations.
\item $X_{i4}$ = annual mean fuel of test locations.
\item $X_{i5}$ = annual mean wind of test locations.
\end{enumerate}

$\epsilon \sim \mathcal{N}(0, \sigma^2)$ denotes the random variation with constant variance.


Conducting F- Test


Hypotheses: H0: B1 = B2 = B3 = B4 = B5 = 0 vs H1: Not all Bi = 0, i = 1,2,..5


Hypotheses:
$$H_0: B_1 = B_2 = B_3 = B_4 = B_5 = 0 $$
$$H_1: \text{Not all } B_i = 0, \text{ where } i = 1,2,\ldots,5$$

```{r}

anova(Model1)

```




\begin{tabular}{|l|c|c|c|c|c|}
\hline
 & Df. & Sum sq. & Mean sq. & F value & Pr(>F) \\
\hline
Regression & 5 & 15703.5 & 3140.7 & 31.95015 & 0 \\
Residuals & 56 & 5502.6 & 98.3 & & \\
\hline
\end{tabular}





  * Note: Regression SS = 4742.6 + 1992.7 + 8651.9 + 258.1 + 58.2.,

  * Therefore, Mean sq reg = Reg SS/Reg Df = 15703.5/5. 
 
 
  * Test statistic: $F_{\text{obs}} = \frac{\text{MS}_{\text{reg}}}{\text{MS}_{\text{Res}}} = \frac{3140.7}{98.3} \approx 31.95015$.


* The null distribution for the test statistic is $F_{5,56}$.
 
 
* P-value: $P(F_{5,56} \geq 31.95015) = 0$.
 
### d) For the Diagnostics
```{r}

  par(mfrow = c(3, 2))
plot(Model1,which = 1:2)
plot(resid(Model1)~transport, data = traffic, xlab="transport",ylab = "Residuals")
abline(h=0,lty=2)
plot(resid(Model1)~road, data = traffic, xlab="road", ylab="Residuals")
abline(h=0,lty=2)
plot(resid(Model1)~weather, data = traffic, xlab = "weather", ylab = "Resdiuals")
abline(h=0,lty=2)
plot(resid(Model1)~fuel, data = traffic, xlab="fuel", ylab = "Residuals")
abline(h=0,lty=2)
plot(resid(Model1)~wind, data = traffic, xlab="wind", ylab = "Residuals")
abline(h=0,lty=2)
```


*The residuals' quantile plot is roughly linear, indicating that the normalcy 
assumption for residuals is valid. All of the residual plots show no obvious trend, 
suggesting that the multiple linear model's constant variance and linearity assumptions are supported. 

  
### e) Finidng R square 
```{r}

summary(Model1)
```



*Here, the goodness of fit metric is R square, which equals 0.7405 = 74.05%.
This indicates that the entire linear regression model explains 74.05% of 
the variation in spi.


